---
title: "Abortion Data Exporation"
author: "Avery Chan"
date: "10/10/2021"
output:
  html_document:
    toc: true
    df_print: paged
    code_folding: show
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggplot2)
library(ggprism)
theme_set(theme_bw())
knitr::opts_chunk$set(rows.print=20)
```

## Description of Data by Source ([stevedata](http://svmiller.com/stevedata/reference/gss_abortion.html))

> This is a toy data set derived from the General Social Survey that I intend to use for several purposes. First, the battery of abortion items can serve as toy data to illustrate mixed effects modeling as equivalent to a one-parameter (Rasch) model. Second, I include some covariates to also do some basic regressions. I think abortion opinions are useful learning tools for statistical inference for college students. Third, there's a time-series component as well for understanding how abortion attitudes have changed over time.  
>
> Data include all General Social Survey observations from 1972 to 2018 for these variables. Be mindful of missing data.

| Variable | Description |
|---|---|
| id | a unique respondent identifier |
| year | the survey year |
| age | the respondent's age in years |
| race | the respondent's race, as character variable |
| sex | the respondent's gender, as character variable |
| hispaniccat | the respondent's Hispanic ethnicity, as character variable |
| educ | how many years the respondent spent in school |
| partyid | the respondent's party identification, as character variable |
| relactiv | the self-reported religious activity of the respondent on a 1:11 scale |
| abany | a binary variable that equals 1 if the respondent thinks abortion should be legal for any reason. 0 indicates no support for abortion for any reason. |
| abdefect | a numeric vector that equals 1 if the respondent thinks abortion should be legal if there is a serious defect in the fetus. 0 indicates no support for abortion in this circumstance. |
| abnomore | a numeric vector that equals 1 if the respondent thinks abortion should be legal if a woman is pregnant but wants no more children. 0 indicates no support for abortion in this circumstance. |
| abhlth | a numeric vector that equals 1 if the respondent thinks abortion should be legal if a pregnant woman's health is in danger. 0 indicates no support for abortion in this circumstance. |
| abpoor | a numeric vector that equals 1 if the respondent thinks abortion should be legal if a pregnant woman is poor and cannot afford more children. 0 indicates no support for abortion in this circumstance. |
| abrape | a numeric vector that equals 1 if the respondent thinks abortion should be legal if the woman became pregnant because of a rape. 0 indicates no support for abortion in this circumstance. |
| absingle | a numeric vector that equals 1 if the respondent thinks abortion should be legal if a pregnant woman is single and does not want to marry the man who impregnated her. 0 indicates no support for abortion in this circumstance. |
| pid | `partyid` recoded so that 7 = NA |
| hispanic | a dummy variable that equals 1 if the respondent is any way Hispanic |

## Import Data

```{r}
# Data from http://svmiller.com/stevedata/reference/gss_abortion.html
raw_abr <- read_csv(
  "data/gss_abortion.csv",
  col_types = cols(
    # row_num = col_integer(),
    # id = col_integer(),
    # year = col_integer(),
    # age = col_integer(),
    # race = col_character(),
    # sex = col_character(),
    # hispaniccat = col_character(),
    # educ = col_integer(),
    # partyid = col_character(),
    # relactiv = col_integer(),
    # abany = col_logical(),
    # abdefect = col_logical(),
    # abnomore = col_logical(),
    # abhlth = col_logical(),
    # abpoor = col_logical(),
    # abrape = col_logical(),
    # absingle = col_logical(),
    # pid = col_integer(),
    # hispanic = col_logical()
    row_num = col_integer(),
    id = col_integer(),
    year = col_integer(),
    age = col_integer(),
    race = col_character(),
    sex = col_character(),
    hispaniccat = col_character(),
    educ = col_integer(),
    partyid = col_character(),
    relactiv = col_integer(),
    abany = col_integer(),
    abdefect = col_integer(),
    abnomore = col_integer(),
    abhlth = col_integer(),
    abpoor = col_integer(),
    abrape = col_integer(),
    absingle = col_integer(),
    pid = col_integer(),
    hispanic = col_integer()
  )
) %>% rename(
  row_number = row_num,
  id = id,
  year = year,
  age = age,
  race = race,
  sex = sex,
  hispanic_ethnicity = hispaniccat, # I drop this row as well because I do not know what the numbers mean
  education = educ,
  party_id = partyid,
  religious_activity = relactiv,
  ab_any = abany,
  ab_defect = abdefect,
  ab_no_more = abnomore,
  ab_health = abhlth,
  ab_poor = abpoor,
  ab_rape = abrape,
  ab_single = absingle,
  party_id_na = pid, # recoded party_id, I remove this later
  hispanic = hispanic
)

raw_abr$party_id_na <- NULL
raw_abr$hispanic_ethnicity <- NULL
```

<!-- Getting parsing error for the `relactiv` column which is "the self-reported religious activity of the respondent on a 1:11 scale." It expects it to be binary but it is actually on a 1:11 scale. -->

### Quick View

```{r}
raw_abr
# raw_abr %>% dim()
# raw_abr %>% view()
raw_abr %>% summary()
# raw_abr %>% str()
```

### Missing Values

```{r}
n = raw_abr %>% nrow()
na_counts <- raw_abr %>% is.na() %>% colSums() %>% as_tibble(rownames = "rowname") %>% mutate(p = value / n)
na_counts %>% arrange(desc(p))
```

## Filtering

```{r}
f_abr <-
  raw_abr %>%
  drop_na(
    party_id,
    education,
    age,
    ab_any,
    ab_rape,
    ab_defect,
    ab_no_more,
    ab_health,
    ab_poor,
    ab_single
  )
```

Replace NA values
```{r}
f_abr$hispanic[is.na(f_abr$hispanic)] <- 2
# f_abr$ab_any[is.na(f_abr$ab_any)] <- 2
# f_abr$ab_rape[is.na(f_abr$ab_rape)] <- 2
# f_abr$ab_defect[is.na(f_abr$ab_defect)] <- 2
# f_abr$ab_no_more[is.na(f_abr$ab_no_more)] <- 2
# f_abr$ab_health[is.na(f_abr$ab_health)] <- 2
# f_abr$ab_poor[is.na(f_abr$ab_poor)] <- 2
# f_abr$ab_single[is.na(f_abr$ab_single)] <- 2
# f_abr$religious_activity[is.na(f_abr$religious_activity)] <- 2
```

```{r}
n = raw_abr %>% nrow()
na_counts_raw <-
  raw_abr %>% is.na() %>% colSums() %>% as_tibble(rownames = "rowname") %>% mutate(p = value / n) %>% arrange(desc(p))

na_counts_raw %>% ggplot(aes(x = reorder(rowname, -value), y = value)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  geom_text(aes(label = value), vjust = -0.3, size = 3.5) +
  theme(axis.text.x = element_text(
    angle = 90,
    vjust = 0.5,
    hjust = 1
  )) +
  labs(title = paste("Missing Values (n=", n, ")"),
       x = "Variable",
       y = "Number of missing values")

n = f_abr %>% nrow()
na_counts <-
  f_abr %>% is.na() %>% colSums() %>% as_tibble(rownames = "rowname") %>% mutate(p = value / n)
na_counts <- na_counts %>% arrange(desc(p))

na_counts %>% ggplot(aes(x = reorder(rowname, -value), y = value)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  geom_text(aes(label = value), vjust = -0.3, size = 3.5) +
  theme(axis.text.x = element_text(
    angle = 90,
    vjust = 0.5,
    hjust = 1
  )) +
  labs(title = paste("Missing Values (n=", n, ")"),
       x = "Variable",
       y = "Number of missing values")
```

## Replace NA values for plotting

```{r}
# # f_abr$year %>% unique()
# missing_religion <- raw_abr %>% 
#   group_by(year) %>% 
#   summarise(
#     num_na = sum(is.na(ab_rape)),
#     n = n(),
#     # had_question = !(num_na == n),
#     p = num_na / n,
#     had_question = p < 0.9,
#   )
# 
# missing_religion %>% ggplot(aes(x = year, y = n, fill=had_question)) +
#   geom_bar(stat = "identity") +
#   # geom_text(aes(label = p), vjust = -0.3, size = 3.5) +
#   theme(axis.text.x = element_text(
#     angle = 90,
#     vjust = 0.5,
#     hjust = 1
#   )) +
#   labs(title = paste("Missing Values for X Per Year"),
#        x = "Year",
#        y = "Number of missing values")

missing_religion <- raw_abr %>%
  group_by(year) %>%
  summarise(
    num_na = sum(is.na(ab_rape)),
    n = n(),
    # had_question = !(num_na == n),
    p = num_na / n,
    had_question = p < 0.9,
  )
```

```{r}
graph_missing <- function(col) {
  missing_x <- raw_abr %>%
    group_by(year) %>%
    summarise(
      num_na = sum(is.na(!!rlang::sym(col))),
      n = n(),
      # had_question = !(num_na == n),
      p = num_na / n,
      had_question = p < 0.9,
    )
  
  missing_x %>% ggplot(aes(x = year, y = n, fill = p)) +
    geom_bar(stat = "identity") +
    geom_text(aes(label = round(p, 2)), vjust = -0.3, size = 1.5) +
    theme(axis.text.x = element_text(
      angle = 90,
      vjust = 0.5,
      hjust = 1
    )) +
    labs(title = paste(paste("Missing Values for ", col, " Per Year")),
         x = "Year",
         y = "Number of missing values")
}
graph_missing("religious_activity")
graph_missing("ab_any")
graph_missing("ab_rape")
graph_missing("ab_defect")
graph_missing("ab_no_more")
graph_missing("ab_health")
graph_missing("ab_poor")
graph_missing("ab_single")
```
```{r}
f_abr$party_id %>% unique()

f_abr <- f_abr %>% mutate(
  party_val = case_when(
    party_id == "Strong Democrat" ~ -3,
    party_id == "Not Str Democrat" ~ -2,
    party_id == "Ind,Near Dem" ~ -1,
    party_id == "Independent" ~ 0,
    party_id == "Ind,Near Rep" ~ 1,
    party_id == "Not Str Republican" ~ 2,
    party_id == "Strong Republican" ~ 3,
    party_id == "Other Party" ~ 0,
  )
)
```


```{r}
f_abr$religious_activity %>% unique()
f_abr$hispanic %>% unique()
write_csv(f_abr, "data/mutated_gss_abortion.csv")
```





















