---
title: "Abortion Data Exporation"
author: "Avery Chan"
date: "10/10/2021"
output:
  html_document:
    toc: true
    df_print: paged
    code_folding: show
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggplot2)
library(ggprism)
theme_set(theme_bw())
knitr::opts_chunk$set(rows.print=20)
```

## Description of Data by Source ([stevedata](http://svmiller.com/stevedata/reference/gss_abortion.html))

> This is a toy data set derived from the General Social Survey that I intend to use for several purposes. First, the battery of abortion items can serve as toy data to illustrate mixed effects modeling as equivalent to a one-parameter (Rasch) model. Second, I include some covariates to also do some basic regressions. I think abortion opinions are useful learning tools for statistical inference for college students. Third, there's a time-series component as well for understanding how abortion attitudes have changed over time.  
>
> Data include all General Social Survey observations from 1972 to 2018 for these variables. Be mindful of missing data.

| Variable | Description |
|---|---|
| id | a unique respondent identifier |
| year | the survey year |
| age | the respondent's age in years |
| race | the respondent's race, as character variable |
| sex | the respondent's gender, as character variable |
| hispaniccat | the respondent's Hispanic ethnicity, as character variable |
| educ | how many years the respondent spent in school |
| partyid | the respondent's party identification, as character variable |
| relactiv | the self-reported religious activity of the respondent on a 1:11 scale |
| abany | a binary variable that equals 1 if the respondent thinks abortion should be legal for any reason. 0 indicates no support for abortion for any reason. |
| abdefect | a numeric vector that equals 1 if the respondent thinks abortion should be legal if there is a serious defect in the fetus. 0 indicates no support for abortion in this circumstance. |
| abnomore | a numeric vector that equals 1 if the respondent thinks abortion should be legal if a woman is pregnant but wants no more children. 0 indicates no support for abortion in this circumstance. |
| abhlth | a numeric vector that equals 1 if the respondent thinks abortion should be legal if a pregnant woman's health is in danger. 0 indicates no support for abortion in this circumstance. |
| abpoor | a numeric vector that equals 1 if the respondent thinks abortion should be legal if a pregnant woman is poor and cannot afford more children. 0 indicates no support for abortion in this circumstance. |
| abrape | a numeric vector that equals 1 if the respondent thinks abortion should be legal if the woman became pregnant because of a rape. 0 indicates no support for abortion in this circumstance. |
| absingle | a numeric vector that equals 1 if the respondent thinks abortion should be legal if a pregnant woman is single and does not want to marry the man who impregnated her. 0 indicates no support for abortion in this circumstance. |
| pid | `partyid` recoded so that 7 = NA |
| hispanic | a dummy variable that equals 1 if the respondent is any way Hispanic |

## Import Data

```{r}
# Data from http://svmiller.com/stevedata/reference/gss_abortion.html
raw_abr <- read_csv(
  "gss_abortion.csv",
  col_types = cols(
    row_num = col_integer(),
    id = col_integer(),
    year = col_integer(),
    age = col_integer(),
    race = col_character(),
    sex = col_character(),
    hispaniccat = col_character(),
    educ = col_integer(),
    partyid = col_character(),
    relactiv = col_integer(),
    abany = col_logical(),
    abdefect = col_logical(),
    abnomore = col_logical(),
    abhlth = col_logical(),
    abpoor = col_logical(),
    abrape = col_logical(),
    absingle = col_logical(),
    pid = col_integer(),
    hispanic = col_logical()
  )
) %>% rename(
  row_number = row_num,
  id = id,
  year = year,
  age = age,
  race = race,
  sex = sex,
  hispanic_ethnicity = hispaniccat, # I drop this row as well because I do not know what the numbers mean
  education = educ,
  party_id = partyid,
  religious_activity = relactiv,
  ab_any = abany,
  ab_defect = abdefect,
  ab_no_more = abnomore,
  ab_health = abhlth,
  ab_poor = abpoor,
  ab_rape = abrape,
  ab_single = absingle,
  party_id_na = pid, # recoded party_id, I remove this later
  hispanic = hispanic
)

raw_abr$party_id_na <- NULL
raw_abr$hispanic_ethnicity <- NULL
```

<!-- Getting parsing error for the `relactiv` column which is "the self-reported religious activity of the respondent on a 1:11 scale." It expects it to be binary but it is actually on a 1:11 scale. -->

### Quick View

```{r}
raw_abr
# raw_abr %>% dim()
# raw_abr %>% view()
raw_abr %>% summary()
# raw_abr %>% str()
```

### Missing Values

```{r}
n = raw_abr %>% nrow()
na_counts <- raw_abr %>% is.na() %>% colSums() %>% as_tibble(rownames = "rowname") %>% mutate(p = value / n)
na_counts %>% arrange(desc(p))
```

## Filtering

```{r}
f_abr <-
  raw_abr %>% filter(!is.na(ab_any)) # drop rows where general stance on abortion wasn't given

n = raw_abr %>% nrow()
na_counts_raw <-
  raw_abr %>% is.na() %>% colSums() %>% as_tibble(rownames = "rowname") %>% mutate(p = value / n) %>% arrange(desc(p))

na_counts_raw %>% ggplot(aes(x = reorder(rowname, -value), y = value)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  geom_text(aes(label = value), vjust = -0.3, size = 3.5) +
  theme(axis.text.x = element_text(
    angle = 90,
    vjust = 0.5,
    hjust = 1
  )) +
  labs(title = paste("Missing Values (n=", n, ")"),
       x = "Variable",
       y = "Number of missing values")

n = f_abr %>% nrow()
na_counts <-
  f_abr %>% is.na() %>% colSums() %>% as_tibble(rownames = "rowname") %>% mutate(p = value / n)
na_counts <- na_counts %>% arrange(desc(p))

na_counts %>% ggplot(aes(x = reorder(rowname, -value), y = value)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  geom_text(aes(label = value), vjust = -0.3, size = 3.5) +
  theme(axis.text.x = element_text(
    angle = 90,
    vjust = 0.5,
    hjust = 1
  )) +
  labs(title = paste("Missing Values (n=", n, ")"),
       x = "Variable",
       y = "Number of missing values")
```

## Replace NA values for plotting

```{r}
f_abr
write_csv(f_abr, "mutated_gss_abortion.csv")
```





















